An integrated distance sampling approach for modelling age-structured populations

### 2023-06-02 Manuscript_Nilsen_Nater.qmd

-   Distance sampling -- such methods are still lacking/not well utilized

-   Integrated pop. Model;

@@

Distance sampling models have been used for estimating animal abundance in a wide range of contexts and for a variety of taxa (+ref). In recent years, the framework has been extended and also applied for estimating changes in abundance over time in open populations by embedding a hidden state model representing population dynamics [@moore2011; @sollmann2015].

>>>>

Distance sampling (DS) models have been used for estimating animal abundance in a wide range of contexts and for a variety of taxa [@buckland2015]. In recent years, they have been extended and also applied for estimating changes in abundance over time in open populations by embedding a hidden state model representing population dynamics [@moore2011; @sollmann2015]. These latter frameworks may allow to accurately quantify population changes but provide little information on the drivers of these changes, i.e. the underlying vital rates. 
Concurrent with the development of more sophisticated DS models, another group of models has emerged and gained popularity, not least for their ability to disentangle demographic processes underlying population dynamic: integrated population models [IPMs, @schaub2021]. Through joint analysis of multiple datasets, IPMs allow simultaneous estimation of population size and composition, as well as all vital rates that form part of an underlying typically age- or stage-structured population model. 
Since both DS models and IPMs estimate population size/density, a combination of the two frameworks has the potential to provide good estimates of both population- and demographic parameters and maximizing knowledge gained through transect surveys by augmenting them with other available data [e.g. @schmidt2020]. 
In this study [...]

@@
